#  Load Data {#load-data}


A search on the GEO web site^[this search was conducted in 2017]
identified the following datasets as potentialy useful for this 
exercise.  All datasets have Affymetrix gene expression data for a 
number of samples annotated with msi status:

- GSE4554 [@Watanabe:2006aa]: [HG-U133_Plus_2] (84 CRC = 33 MSI + 51 MSS)
- GSE13067 [@Jorissen:2008aa]: [HG-U133_Plus_2] (74 CRC)
- GSE13294 [@Jorissen:2008aa]:  [HG-U133_Plus_2] (89 MSI + 140 MSS + 58 MSI + 77 MSS)
- GSE24514 [@Alhopuro:2012aa]: [HGU133A] (34 MSI,15 N)
- GSE30540 [@Watanabe:2012aa]: [HG-U133_Plus_2] (35 stage II and stage III)
- GSE35896 [@Schlicker:2012aa]: [HG-U133_Plus_2] (62 CRC samples)
- GSE75316 [@Barras:2017aa]: [HG-U133_Plus_2] (59 CRC samples)
- GSE39084 [@Kirzin:2014aa]:  [HG-U133_Plus_2] (70 CRC samples)
- GSE26682 [@Schmit:2015aa]: [HG-U133A], [HG-U133_Plus_2] (???)
- GSE14526 [@Yagi:2010aa]: [HG-U133_Plus_2]
- GSE4045 [@Laiho:2006aa]: [HG-U133A] 

## Download Datasets 

This process is somewhat manual as the GEO data do not follow an entirely standard
format.  Each dataset will have to be downloaded one at a time and examined.
Following this step, we can loop again through the datasets and store the necessary
data in a uniform manner.


We can download data from GEO using function 
`getGEO` from package `GEOquery`.


```{r load-data-dnldGSE, eval=FALSE, cache=T, cache.vars='',eval=F,message=F}

 suppressMessages(require(GEOquery))

 GSE_SET.vec <- c('GSE4554', 'GSE13067', 'GSE13294', 
                  'GSE24514', 'GSE30540', 'GSE35896',
                  'GSE75316', 'GSE39084', 'GSE26682',
                  'GSE14526', 'GSE4045')

 for(SET in GSE_SET.vec){
  DnLd.tm <- startTimedMessage("Start of Download for ", SET)
   Set.gse <- getGEO(SET, destdir=temp_DIR ,getGPL=F) 
   saveObj(paste0(SET,'.gse'), 'Set.gse')
  stopTimedMessage(DnLd.tm)

 }

```
<!-- ######################################################################## -->

## Peek into Datasets

```{r load-data-peekGSE, cache=TRUE, cache.vars='', fig.width=11, fig.height=6,eval=T}
 # CHANGE THIS LINE TO CLEAR CACHE 
 suppressMessages(require(Biobase))
 gseFile.vec <- list.files(file.path('RData'), 'gse$')

gsePhenoData.lst <- list()
gseDim.lst <- list()

 for(gseF in gseFile.vec){
  cat("\nPeek at ", gseF, '\n')

  loadObj(gseF, 'Set.gse')

gseName <- sapply(strsplit(gseF, split='\\.'),'[',1)

  print(dim(exprs(Set.gse[[1]])))
  if(nrow(exprs(Set.gse[[1]])) == 0) next()  

gseDim.lst[[gseName]] <- dim(exprs(Set.gse[[1]]))

 #################################
 # Expression Summary 
if(F) {#SKIP
  cat("\nExpression Summary:\n")
  knitr::kable(t(apply(exprs(Set.gse[[1]]),2,summary))) %>%
      kableExtra::kable_styling(full_width = F)

}@SKIP

 #################################
 # Expression Boxplots
 old_par <- par(mar=par('mar')+c(2,0,0,0))

 boxplot(exprs(Set.gse[[1]]), outline=F, las=2) 
 title(paste("Boxplots of expression values for", gseF))

 par(old_par)

 
 #################################
gsePhenoData.lst[[gseName]] <- phenoData(Set.gse[[1]])@data

if(F) {#SKIP
 cat("\nSample Descriptions:\n")
 print(knitr::kable(phenoData(Set.gse[[1]])@data) %>%
        kableExtra::kable_styling(full_width = F))
 }
}#SKIP

```
<!-- ######################################################################## -->
  

## Dataset Cleaning

Following a peek into the datasets, we will proceed forward keeping only the
datasets which used the HG-U133_Plus_2 gene chip (N=54675 probe sets) and have
msi status:  
<!-- Reuires combing through gsePhenoData.lst names -->
* `r KEEP_GSE.vec <- c('GSE13067', 'GSE13294','GSE35896','GSE39084', 'GSE4554');  KEEP_GSE.vec`

We exclude GSE75316 from the analysis as most of the samples are already part of GSE13067 and the latter
is a larger set.  

In the next chunk, we go through each of these datasets and store the gene expression and 
sample description data in a uniform manner to facilitate subsequent pooling.  We also
take a look at potential intra dataset batch effects. 

```{r load-data-cleanGSE, echo=T, cache=TRUE, cache.vars='', fig.width=8, fig.height=6,eval=T}
 # CHANGE THIS LINE TO CLEAR CACHE ..
 suppressMessages(require(hgu133plus2.db))

 for(SET in KEEP_GSE.vec){
  cat("\n\nProcessing ", SET, '\n')

  loadObj(paste0(SET, '.gse'), 'Set.gse')

  ###################################################################
  #  Expr - matrix
  #  - rows are features
  #  - columns are samples
  ###################################################################
  Expr.mtx <- exprs(Set.gse[[1]])

  if(SET %in% c('GSE4554', 'GSE4045'))  Expr.mtx <- log2(pmax(Expr.mtx, min(Expr.mtx[Expr.mtx>0])))

  ###################################################################
  #  SampAttr - data.frame
  #  - rows are samples
  #  - columns are atttibutes of the samples
  ###################################################################
  phenoData.frm <- phenoData(Set.gse[[1]])@data

  if(nrow(phenoData.frm) != ncol(Expr.mtx)) stop("Expression-sampAttr Mismatch for ", SET)

  MS_Status <- switch(SET,
      GSE13067 = gsub(' ','', 
                 sapply(strsplit(phenoData.frm$title, split=':'), function(x) x[2])), 
      GSE13294 = gsub(' ','', 
                 sapply(strsplit(phenoData.frm$title, split=':'), function(x) x[2])), 
      GSE35896 = gsub(' ','', 
                 sapply(strsplit(phenoData.frm$characteristics_ch1.7, split=':'), function(x) x[2])), 
      GSE39084 = gsub(' ','', 
                 sapply(strsplit(phenoData.frm$characteristics_ch1.16, split=':'), function(x) x[2])), 
      GSE4554 =  gsub(' ', '',
                 sapply(strsplit(as.character(phenoData.frm$description), split=','), function(x) x[1])),
        "ERROR")


  KEEP.ndx <- which(is.element(MS_Status, c('MSS', 'MSI', 'No', 'Low', 'High')))

  sampAttr.frm <- data.frame(Sample_Id=rownames(phenoData.frm),
                             MS_Status=MS_Status, 
                             row.names=rownames(phenoData.frm))[KEEP.ndx,]
  Expr.mtx <- Expr.mtx[,rownames(sampAttr.frm)]
 
  print(table(MS_Status[KEEP.ndx]))

  ###################################################################
  #  featureAttr - data.frame
  #  - rows of featureAttr match rows of Expr mtx
  #  - columns of featureAttr are atttibutes of the features - eg geneSymbol
  ###################################################################
  # This deosn;t work!
  #featureAttr.frm <- featureData(Set.gse[[1]])@data
  #dim(featureAttr.frm)
  #Symbol.vec <- mget(featureAttr.frm$ID, hgu133plus2SYMBOL)
  #featureAttr.frm$Symbol <- Symbol.vec[rownames(featureAttr.frm)]
 
  Symbol.vec <- mget(rownames(Expr.mtx), hgu133plus2SYMBOL)
  featureAttr.frm <- data.frame(Symbol=unlist(Symbol.vec))
  rownames(featureAttr.frm) <- rownames(Expr.mtx)
 
  ###################################################
  # Save in arrayGeneExpr object
  ###################################################
  if(sum(colnames(Expr.mtx) != rownames(sampAttr.frm))) stop("Column order error")
  if(sum(rownames(Expr.mtx) != rownames(featureAttr.frm))) stop("Row order error")
 
  Set.arrayGeneExpr <-  list(exprData=NA, exprType=NA, Norm=NA,
                                   sampAttr=NA, featureAttr=NA)
  class(Set.arrayGeneExpr) <- "arrayGeneExpr"
  Set.arrayGeneExpr$exprData <- Expr.mtx
  Set.arrayGeneExpr$sampAttr <- sampAttr.frm
  Set.arrayGeneExpr$featureAttr <- featureAttr.frm
 
  saveObj(paste0(SET,'.arrayGeneExpr'), 'Set.arrayGeneExpr')
 
  ######################################################
  # Look for intra-set batch effects
  ######################################################
  Expt.prcomp <- prcomp(Expr.mtx)
  
  plot(x=Expt.prcomp$rotation[,'PC1'],
       xlab=paste0('PC1 (', round(100*summary(Expt.prcomp)$importance[2,"PC1"],1),'%)'),
       y=Expt.prcomp$rotation[,'PC2'],
       ylab=paste0('PC2 (', round(100*summary(Expt.prcomp)$importance[2,"PC2"],1),'%)'),
       pch=19, cex=1.5, col=as.numeric(as.factor(sampAttr.frm$MS_Status)))
  title(SET) 
  legend(ifelse(SET %in% c('GSE13067', 'GSE35896'), 'bottomleft', 'bottomright'),
  legend=levels(as.factor(sampAttr.frm$MS_Status)),
  col=1:length(levels(as.factor(sampAttr.frm$MS_Status))),
  pch=19)
 
 } 

```
<!-- ######################################################################## -->

## Data Pooling


When pooling data from different sources, we should make sure not to introduce batch effects
into the mix.  In the first instance we will simply pool the data sets together,
and apply quantile normalization.  This can be seen as a very weak form of batch effect
correction and is a minimal requirement.  For a more careful way to remove batch effects
and other unwanted variation see
Gagnon-Bartsch (2012) [@Gagnon-Bartsch:2012aa] and 
Molania et al. (2019) [@Molania:2019aa].


```{r load-data-poolGSE,cache=TRUE,cache.vars='',fig.width=12, fig.height=6,eval=T}

 ###############################
 # Get feature attributes from one of the arrayGeneExpr objects
 ###############################
 loadObj(paste0(KEEP_GSE.vec[1],'.arrayGeneExpr'), 'Set.arrayGeneExpr')

 featureAttr.frm <- Set.arrayGeneExpr$featureAttr
 featureAttr.frm$Symbol[is.na(featureAttr.frm$Symbol)] <-
        rownames(featureAttr.frm)[is.na(featureAttr.frm$Symbol)]
 featureAttr.frm[1:5,,drop=F]

 geneId.vec <- rownames(featureAttr.frm)

 ###############################

 ###############################
 # Pool the expression data
 ###############################
 PooledExpr.mtx <- do.call('cbind', lapply(KEEP_GSE.vec,
  function(SET) {
    loadObj(paste0(SET,'.arrayGeneExpr'), 'Set.arrayGeneExpr')
    Set.arrayGeneExpr$exprData[geneId.vec,]
 }))

 #dim(PooledExpr.mtx)
 #PooledExpr.mtx[1:5, 1:5]

 # Apply quantile normalization
 median.Expr.vec <- apply(apply(PooledExpr.mtx,2,sort), 1, median)
 #summary(median.Expr.vec)

 PooledNormedExpr.mtx <- apply(PooledExpr.mtx, 2, function(x)
     median.Expr.vec[rank(x)])
 rownames(PooledNormedExpr.mtx) <- rownames(PooledExpr.mtx)


 #summary(apply(PooledExpr.mtx, 2, mean))
 #summary(apply(PooledNormedExpr.mtx, 2, mean))

 ###############################
 # Pool the sample attribute data
 ###############################
 PooledSampAttr.frm <- do.call('rbind', lapply(KEEP_GSE.vec,
  function(SET) {
   #cat(SET,'\n')
    loadObj(paste0(SET,'.arrayGeneExpr'), 'Set.arrayGeneExpr')
    data.frame(GSE=rep(SET,nrow(Set.arrayGeneExpr$sampAttr)),
               Sample_Id=Set.arrayGeneExpr$sampAttr$Sample_Id,
               MS_Status=Set.arrayGeneExpr$sampAttr$MS_Status,
               row.names=rownames(Set.arrayGeneExpr$sampAttr))
  }))

 #dim(PooledSampAttr.frm)
 #PooledSampAttr.frm[1:5, ]

 # Assemble arrayGeneExpr object for pooled data and save
 GSEPool.arrayGeneExpr <-  list(exprData=NA, exprType=NA, Norm=NA,
                                  sampAttr=NA, featureAttr=NA)
 class(GSEPool.arrayGeneExpr) <- "arrayGeneExpr"
 GSEPool.arrayGeneExpr$exprData <- PooledNormedExpr.mtx
 GSEPool.arrayGeneExpr$exprType <- 'MAS5_RMA'
 GSEPool.arrayGeneExpr$Norm <- 'Quantile'
 GSEPool.arrayGeneExpr$sampAttr <- PooledSampAttr.frm
 GSEPool.arrayGeneExpr$featureAttr <- featureAttr.frm

 save(GSEPool.arrayGeneExpr, file=file.path('RData','GSEPool.arrayGeneExpr'))

 ######################################################
 # Look for inter-set batch effects
 ######################################################
 Expt.prcomp <- prcomp(PooledNormedExpr.mtx)
  
 par(mfrow=c(1,2), mar=c(3,3,2,1))
 # Annotate with Batch
 plot(x=Expt.prcomp$rotation[,'PC1'],
      xlab=paste0('PC1 (', round(100*summary(Expt.prcomp)$importance[2,"PC1"],1),'%)'),
      y=Expt.prcomp$rotation[,'PC2'],
      ylab=paste0('PC2 (', round(100*summary(Expt.prcomp)$importance[2,"PC2"],1),'%)'),
      pch=19, cex=1.5, col=as.numeric(as.factor(PooledSampAttr.frm$GSE)))
 title('Color is Batch')

 legend.tbl <- table(as.factor(PooledSampAttr.frm$GSE), 
                     as.numeric(as.factor(PooledSampAttr.frm$GSE)))
 legend('topright', pch=19, cex=1.5, 
         legend=rownames(legend.tbl),
         col=as.numeric(colnames(legend.tbl)))

 # Annotate with MS_Status
 MSS_MSI <- as.factor(ifelse(is.element(PooledSampAttr.frm$MS_Status, c('MSS', 'No')),
          'MSS','MSI'))
 plot(x=Expt.prcomp$rotation[,'PC1'],
      xlab=paste0('PC1 (', round(100*summary(Expt.prcomp)$importance[2,"PC1"],1),'%)'),
      y=Expt.prcomp$rotation[,'PC2'],
      ylab=paste0('PC2 (', round(100*summary(Expt.prcomp)$importance[2,"PC2"],1),'%)'),
      pch=19, cex=1.5, col=as.numeric(as.factor(MSS_MSI)))
 title('Color is MSS_MSI')

 legend.tbl <- table(as.factor(MSS_MSI), 
                     as.numeric(as.factor(MSS_MSI)))
 legend('topright', pch=19, cex=1.5, 
         legend=rownames(legend.tbl),
         col=as.numeric(colnames(legend.tbl)))

```
<!-- ######################################################################## -->


## Batch Correction

We see a definite batch effect in the PCA plot based on the pooled expression data.
The clustering is along probe set reduction method, RMA vs MAS 5.0.
One should really go back to the cel files and re-analyze all samples using one
probe set reduction method.  For the sake of illustration, we will instead try to
remove this effect using a batch-effect correction method.

Given that we don't have a good set of control genes as required by the RUV method, we'll
use 'ComBat` from  the 
[SVA](https://www.bioconductor.org/packages/release/bioc/html/sva.html)
package.

Alternatively, we could keep keep the two sets of data, those summarized using RMA
and those summarized using MAS 5.0, separate, alternating using one set for
model selection and fitting and the other for testing.  This would provide
a strong test of the generalizability and robustness of the results.

```{r load-data-svaBatchAdjust, cache=TRUE, cache.vars='',fig.width=12, fig.height=6,eval=T}
 suppressMessages(require(sva))

 load(file=file.path('RData','GSEPool.arrayGeneExpr'))

 Expr.mtx <- GSEPool.arrayGeneExpr$exprData
 sampAttr.frm <- GSEPool.arrayGeneExpr$sampAttr


 # Applying theComBatfunction to adjust for known batches
 # Note: we could here use a binray batch: RMA vs MAS5.
 batch <- as.factor(sampAttr.frm$GSE)

 with(sampAttr.frm, table(GSE, MS_Status))

 MSS <- as.factor(is.element(sampAttr.frm$MS_Status, c('MSS', 'No')))
 modcombat = model.matrix(~ MSS)

 combat_edata = ComBat(dat=Expr.mtx, 
                       batch=batch, 
                       mod=modcombat, 
                       par.prior=TRUE, 
                       prior.plots=FALSE)

 GSEPool.arrayGeneExpr$exprDataBatchAdj <- combat_edata

 save(GSEPool.arrayGeneExpr, file=file.path('RData','GSEPool.arrayGeneExpr'))

 ######################################################
 # Look for inter-set batch effects
 ######################################################
 Expt.prcomp <- prcomp(GSEPool.arrayGeneExpr$exprDataBatchAdj)
  
 par(mfrow=c(1,2), mar=c(3,3,2,1))
 # Annotate with Batch
 plot(x=Expt.prcomp$rotation[,'PC1'],
      xlab=paste0('PC1 (', round(100*summary(Expt.prcomp)$importance[2,"PC1"],1),'%)'),
      y=Expt.prcomp$rotation[,'PC2'],
      ylab=paste0('PC2 (', round(100*summary(Expt.prcomp)$importance[2,"PC2"],1),'%)'),
      pch=19, cex=1.5, col=as.numeric(as.factor(sampAttr.frm$GSE)))
 title('Color is Batch')

 legend.tbl <- table(as.factor(sampAttr.frm$GSE), 
                     as.numeric(as.factor(sampAttr.frm$GSE)))
SKIP <- function() {
 legend('topright', pch=19, cex=1.5, 
         legend=rownames(legend.tbl),
         col=as.numeric(colnames(legend.tbl)))
}#SKIP

 # Annotate with MS_Status
 MSS_MSI <- as.factor(ifelse(is.element(sampAttr.frm$MS_Status, c('MSS', 'No')),
          'MSS','MSI'))
 plot(x=Expt.prcomp$rotation[,'PC1'],
      xlab=paste0('PC1 (', round(100*summary(Expt.prcomp)$importance[2,"PC1"],1),'%)'),
      y=Expt.prcomp$rotation[,'PC2'],
      ylab=paste0('PC2 (', round(100*summary(Expt.prcomp)$importance[2,"PC2"],1),'%)'),
      pch=19, cex=1.5, col=as.numeric(as.factor(MSS_MSI)))
 title('Color is MSS_MSI')

 legend.tbl <- table(as.factor(MSS_MSI), 
                     as.numeric(as.factor(MSS_MSI)))
 legend('topright', pch=19, cex=1.5, 
         legend=rownames(legend.tbl),
         col=as.numeric(colnames(legend.tbl)))

 
```
<!-- ######################################################################## -->

Visual inspection of the effect of the batch correction transformation
is a minimal requirement for verification.  One can do a better job
at quantifying the effect of batch correction, especially of the data set
contains control features and control samples. See Lazar et.al. [@Lazar:2013aa]
for a discussion of batch effect removal assessment.


##  Separate Data Set into Train and Test Subsets

We will separate the data set into **Train** and **Test** subsets here, before
any other filtering or data manipulation.  In particular, since we will
be interested in evaluating the effect of gene or probe set selection,
this selection must be made on the basis of the training subset only.
This is done as a precautionary measure as gene selection based on
a variability filter is is done here is unlikely to have an effect on sample classification
performance.

Take note that the expression matrices that we save here will
be transposed with `genes in columns`.  This is done to accomodate
the data format expected by the `caret` package.

```{r load-data-getTrainTest, cache=TRUE,cache.vars='',eval=T}
 # CHANGE THIS LINE TO CLEAR CACHE
 suppressMessages(require(caret))

 # Load expression  data object
 load(file.path('RData','GSEPool.arrayGeneExpr'))

 # Get expression data matrix 
 Expr.mtx <- t(GSEPool.arrayGeneExpr$exprDataBatchAdj)

 # Use gene names where possible
 colnames(Expr.mtx) <- 
 make.names(GSEPool.arrayGeneExpr$featureAttr[colnames(Expr.mtx),], unique=T)

 # Keep map
 GeneNameMap.vec <- rownames(GSEPool.arrayGeneExpr$exprDataBatchAdj)
 names(GeneNameMap.vec) <- colnames(Expr.mtx)

 #GeneNameMap.vec[1:5]

 save(GeneNameMap.vec, file=file.path('RData', 'GeneNameMap.vec'))

 # Get sample attributes
 sampAttr.frm <- GSEPool.arrayGeneExpr$sampAttr

 DataSource.vec <- sampAttr.frm$GSE
 names(DataSource.vec) <- rownames(sampAttr.frm)

 Label.vec <- ifelse(is.element(sampAttr.frm$MS_Status, c('MSS', 'No')), 'MSS','MSI')
 names(Label.vec) <- rownames(sampAttr.frm)
 rm(GSEPool.arrayGeneExpr)

 # split Into Train and Test
 set.seed(12379)
 inTrain <- createDataPartition(y=Label.vec, p=0.75, list=F)

 Train.Expr.mtx <- Expr.mtx[inTrain,]
 Train.Label.vec <- Label.vec[inTrain]
 Train.DataSource.vec <- DataSource.vec[inTrain]


 Test.Expr.mtx <- Expr.mtx[-inTrain,]
 Test.Label.vec <- Label.vec[-inTrain]
 Test.DataSource.vec <- Label.vec[-inTrain]

 knitr::kable(rbind(Train=dim(Train.Expr.mtx),
             Test=dim(Test.Expr.mtx))) %>%
        kableExtra::kable_styling(full_width = F)


 knitr::kable(rbind(
 Train=table(Train.Label.vec)/length(Train.Label.vec),
 Test=table(Test.Label.vec)/length(Test.Label.vec))) %>%
        kableExtra::kable_styling(full_width = F)



 # Save these
 save(Train.Expr.mtx, file=file.path('RData', 'Train.Expr.mtx'))
 save(Test.Expr.mtx, file=file.path('RData', 'Test.Expr.mtx'))
 save(Train.Label.vec, file=file.path('RData', 'Train.Label.vec'))
 save(Test.Label.vec, file=file.path('RData', 'Test.Label.vec'))
 save(Train.DataSource.vec, file=file.path('RData', 'Train.DataSource.vec'))
 save(Test.DataSource.vec, file=file.path('RData', 'Test.DataSource.vec'))

```
<!-- ######################################################################## -->


## Save Gene Sets

Save gene sets selected by overall variability *in training samples*.
<!-- In the interest of conserving computing resources,--> In the analysis 
that follows we will use the top `r VAR_FILTER`%
most variable genes.  

Note that we select genes based on variability in the entire training
dataset which includes data from different GEO data sets.  This selection
may favor the inclusion of genes which differ across the different
data sets due to technical reasons.  An alternative selection would be to
to select genes based on within data set variability.  This could be
implemented by filtering based on the residual variance of an
ANOVA model `Expr ~ GSE` fitted to the gene expression data.

```{r load-data-saveGeneSets,cache=TRUE,cache.vars='',eval=T}
 # CHANGE THIS LINE TO CLEAR CACHE
 load(file.path('RData', 'Train.Expr.mtx'))

 # Identify and remove low variance columns
 Train.Expr.mad.vec <- apply(Train.Expr.mtx,2,mad)
 #summary(Train.Expr.mad.vec)

 TopVar.cols <- which(Train.Expr.mad.vec > quantile(Train.Expr.mad.vec, prob=(100-VAR_FILTER)/100))
 Train.TopVarGenes.vec <- colnames(Train.Expr.mtx)[TopVar.cols]
 #length(Train.TopVarVarGenes.vec)

 saveObj(paste0('Train.',SelGenes,'.vec'), 'Train.TopVarGenes.vec')

 # also save top 25 and top 50 - Nawh!

```
<!-- ######################################################################## -->

